Title of submission to PLOS journal

### 2019-08-18 docs/manuscript_draft/manuscript_draft.tex
move images

% Use "Eq" instead of "Equation" for equation citations.
\emph{Text based on plos sample manuscript, see
\url{http://journals.plos.org/ploscompbiol/s/latex}}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

A list

\begin{itemize}
\tightlist
\item
  Item 1
\item
  Item 2
\end{itemize}

@@

@@

>>>>

% Use "Eq" instead of "Equation" for equation citations.
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Understanding the mechanisms underlying parental care are critical to
circumventing issues with parent-newborn bonding as well, where ultimate
explanations are obvious, but specific mechanisms remain elusive.

The rock dove (\emph{Columba livia}) is an ideal system to characterize
changes in genetic expression during parental care transitions because:
1) ample genomic resources are available, including a complete annotated
genome assembly (Gillespie et al.~2013; Shapiro et al.~2013) and
methodology concerning reproductive physiology and behavior (Dong et
al.~2012); and 2) rock doves are prolific, year-round breeders that
thrive in captivity, making observation, manipulation, and sampling
highly feasible year-round.

\includegraphics[width=360px]{../../figures/images/PigeonHPGaxis}

Rock doves are socially monogamous and offer bi-parental care, making
inter- and intra-sexual comparisons possible. Birds offer two important
behavioral transition points into parental care: the incubation of eggs
and the caring for chicks. This produces two unique opportunities to
study how the brain transitions into two different suites of parental
care behaviors. Additionally, rock doves exhibit a parental care
strategy analogous to mammals in that they, too, `lactate' to feed their
young (Gillespie et al.~2011, 2012). This lactation, unlike simple
regurgitation of food, consists of the production and sloughing off of
skin cells inside the crop sac of females and males, creating a
protein-rich milk-like substance on which they rear their chicks. Many
functional similarities between rock dove and mammalian lactation exist
concerning the mediation of this event by the hormone prolactin (Dumont
1965). Additionally, like mammalian milk, rock dove milk delivers
essential immunoglobulins and nutritional benefits to young, aiding in
their immune function and development of microbiota (Gillespie et
al.~2012). Thus, because rock doves incubate eggs and exhibit
mammalian-like mediation and function of lactation for young, they have
the potential to serve as a powerful theoretical bridge to understand
the neurobiology of both avian and mammalian transitions into parental
care.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

@@

\includegraphics[width=360px]{../../figures/images/samplingtimepoints}

\hypertarget{elucidate-cause-of-transcription-protein-changes-during-parental-care-transitions}{%
\subsection{Elucidate cause of transcription \& protein changes during
parental care
transitions}\label{elucidate-cause-of-transcription-protein-changes-during-parental-care-transitions}}

Using egg and chick removal and replacement experiments we will
determine if neural transcription and translation during parental care
transitions are based on external sensory information (the
presence/absence of eggs or chicks) or on an internal clock-timing
mechanism (natural biological rhythms unaffected by changes in external
sensory information). As rock doves are known to cross-foster eggs and
chicks, our working hypothesis is that the transition to parental care
behaviors is based on external sensory input, and specific predictions
are explained in conjunction with each manipulation below.

\hypertarget{give-birds-eggs}{%
\subsubsection{Give birds eggs}\label{give-birds-eggs}}

We will place eggs in nests of birds lacking them and sample brains the
following day. Prediction: birds will transition into incubation
behavior, and neural transcription and translation will mirror that of
birds sampled on the day their first egg was laid. This would support
our working hypothesis that sensory perception of the environment drives
these changes. If we do not observe changes in behavior, transcription,
and translation, or they only occur in males, one of many
interpretations is that males respond to environmental cues while
females are cued by an internal clock-timing mechanism regulating
transitions into parental care.

\hypertarget{prolong-incubation}{%
\subsubsection{Prolong incubation}\label{prolong-incubation}}

We will prolong incubation by removing eggs from actively incubated
nests on the day before hatching (Day 17) and immediately replace them
with infertile, `dummy' eggs. We will then sample brains 3 days after
normal hatch time (Day 21). Collaborator Silver and colleagues have
shown sex differences in the endocrine profiles of male and female doves
in response to prolonged incubation, in that prolactin levels are
maintained in incubating females but not males (Ramsey et al.~1985).
Because of this, we predict that if changes in behavior, neural
transcription, and translation occur in females because of sensory
information related to chicks hatching (supporting our working
hypothesis), then females that experience this prolonged incubation will
mirror the changes of incubating birds sampled on Day 17 of incubation.
However, if changes in behavior, transcription and translation are
brought on by internal clock mechanisms rather than external sensory
information, as predicted at this stage in males, bird behavior and
biology will resemble that of birds collected on Day 3 post-hatching.

\hypertarget{extend-time-of-hatching}{%
\subsubsection{Extend time of hatching}\label{extend-time-of-hatching}}

We will prolong incubation by removing eggs from nests undergoing active
incubation on Day 17 (the end of incubation) and immediately replacing
eggs with `dummy' eggs. On Day 21 (3 days post-natural hatch time) we
will offer hatchlings. Doves are known to care for nestlings placed in
their nest that are not their own (Klinghammer and Hess 1964, Hasen
1971). Brains will be sampled the following day. Prediction: Behavior,
transcription and translation will be similar to birds naturally caring
for chick(s) one day post-hatch of the first chick, supporting our
working hypothesis. Alternatively, if behavior, transcription and
translation are similar to birds caring for hatchlings on Day 3
post-hatching, which would have been the normal time course for parental
care if chicks had hatched naturally, this would support regulation by
an internal clock mechanism.

\hypertarget{remove-eggs}{%
\subsubsection{Remove eggs}\label{remove-eggs}}

We will remove eggs during incubation on Day 1 (beginning), Day 9
(middle), or Day 17 (end) and sample brains on the following day. Note:
This is not a repeated measures sampling, and three groups of
male-female pairs will be independently sampled at each time point.
Prediction: Birds will revert to a behavioral, transcriptional and
translational pattern characteristic of the ``Nesting, prior to lay''
sampling point, lending support to our working hypothesis. If birds
maintain a profile characteristic of birds collected at similar time
points whose eggs have not been removed, this lends support to an
internal clock mechanism.

\hypertarget{reduce-time-to-hatching}{%
\subsubsection{Reduce time to hatching}\label{reduce-time-to-hatching}}

We will remove eggs on Day 8 (middle) of incubation and immediately
replace them with newly hatched chicks. We will sample brains on the
following day (Day 9). Prediction: Changes in behavior, transcription
and translation are externally sensory driven, and these will be similar
to birds sampled on `Day 1: 1st egg hatches, supporting our working
hypothesis. Alternatively, if profiles remain similar to birds sampled
on Day 9 that are actively incubating eggs, this would support the
presence of an internal clock mechanism.

\hypertarget{remove-chicks}{%
\subsubsection{Remove chicks}\label{remove-chicks}}

We will remove newly, naturally hatched chick(s) on Day 2: second chick
hatches and sample brains from parents the following day (Day 2 post
(second) egg hatching). Prediction: Behavior, transcription and
translation are sensory driven and will revert to a pattern
characteristic of the `Nesting, prior to lay' point, supporting our
working hypothesis. Alternatively, if an internal clock mechanism is at
play, we would expect to see a pattern similar to birds caring for
chicks on Day 2 post (second) egg hatching.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

This project is a synergistic collaboration between the PI, Rebecca
Calisi-Rodríguez (expertise in avian behavior, parental care and
neurobiology), co-PI, Matthew MacManes (expertise in next-generation
sequencing, transcriptome assembly, and gene expression analyses), and
Collaborator, Rae Silver (expertise in neurobiology, dove behavior, and
decades of successful breeding and maintenance of dove colonies at
Barnard College).

\hypertarget{metadata-and-extras}{%
\section{Metadata and extras}\label{metadata-and-extras}}

@@

\hypertarget{about-the-authors}{%
\subsection{About the authors}\label{about-the-authors}}

Roles, Affiliations, Orcids, competing interests
